Rhyme: A Query Language for Nested Data Structures

### 2024-12-13 paper.tex
finish section 2

Following IR generation, types are assigned to each IR node. The compiler records the types of generators and temporary variables, as well as the types of data sources associated with each generator. This type information is critical for backend code generation, enabling specialized and efficient co-iteration of multiple data sources. For example, if a generator co-iterates over multiple sparse vectors, \rhyme generates optimized code for advancing the sparse vectors efficiently, accounting for their lack of constant-time look-up. A more detailed explanation of these optimizations is provided in Section \ref{sparse}.
\subsection{\rhyme Backend}
\section{Type System}\label{typesystem}
\section{Sparse Tensor Algebra}\label{sparse}
\section{Evaluation}\label{evaluation}
\section{Conclusion}\label{conclusion}
%\newpage
\bibliographystyle{acm}
\bibliography{paper}
\end{document}
\endinput

>>>>

Following IR generation, types are assigned to each IR node. The compiler records the types of generators and temporary variables, as well as the types of data sources associated with each generator. This type information is critical for backend code generation, enabling specialized and efficient co-iteration of multiple data sources. For example, if a generator co-iterates over multiple sparse vectors, \rhyme generates optimized code for advancing the sparse vectors efficiently, accounting for their lack of constant-time look-up. A more detailed explanation of these optimizations is provided in Section \ref{sparse}.
\subsection{\rhyme Backend}
\iffalse
\rhyme IR does not explicitly capture program dependencies, therefore we need to find an optimal program schedule for the query. We need to find an optimal loop schedule while respecting the correctness of program semantics. Since the IR nodes are only connected through dependencies, we need to figure out the order between loops, how loops should be nested and what instructions to put inside each loop. Putting code inside deeper loop nests could cause redundant computation and will even give us incorrect results if the code represents an aggregation. We also face exponential number of different choices of loop ordering and loop nests, each may cause different performance. We also want to maximize loop fusion such that assignments that depend on the same generator can be put inside one loop to reduce memory usage and loop control overhead. To achieve the goal mentioned above, we formalize the dependency constraints that the code generator needs to respect and derive a heuristic-based algorithm that can maximize loop fusion to ensure performance.\par
\rhyme has multiple backends such as JavaScript, C/C++. Therefore, our IR design is language-agonistic, we do not store any language-specific information in the IR. Instead, we only capture the core program relation (data and control dependencies) in the IR and different language backends can reuse the same dependency infrastructure. The IR nodes will also carry the type information to facilitate specialized cod generation. After the optimal program structure is scheduled, \rhyme will translate each IR nodes to the corresponding backend languages. \rhyme will also use the type information to generate the correct type signature and add explicit type cast when necessary.
\fi
The \rhyme IR does not explicitly encode program dependencies, making it necessary to derive an optimal program schedule for executing queries. This involves determining an optimal loop schedule while ensuring the correctness of the program semantics. Since the IR nodes are connected only through dependency relationships, the code generator must establish the order of loops, determine loop nesting, and decide which instructions to place inside each loop. Incorrectly placing code within deeper loop nests can lead to redundant computations or, in cases involving aggregations, even incorrect results.

The challenge is further compounded by the exponential number of possible loop orderings and nesting configurations, each with varying performance characteristics. To improve performance, it is also essential to maximize loop fusion, enabling assignments dependent on the same generator to be placed within a single loop. This minimizes memory usage and loop control overhead.

To address these challenges, we formalize the dependency constraints that the code generator must respect and design a heuristic-based algorithm that prioritizes loop fusion. This approach ensures both performance optimization and semantic correctness.

\rhyme supports multiple backends, including JavaScript and C/C++. To maintain flexibility, the IR is designed to be language-agnostic and excludes any language-specific details. Instead, it captures only the core program relationships—data and control dependencies. This allows different backends to reuse the same dependency-based IR infrastructure. Each IR node also carries type information to facilitate specialized code generation.

Once an optimal program structure is determined, \rhyme translates the IR nodes into code for the target backend language. During this translation, the type information is used to generate accurate type signatures and insert explicit type casts where necessary. This ensures the generated code is both correct and efficient for the specific backend.\par
\section{Type System}\label{typesystem}
\section{Sparse Tensor Algebra}\label{sparse}
\section{Evaluation}\label{evaluation}
\section{Conclusion}\label{conclusion}
%\newpage
\bibliographystyle{acm}
\bibliography{paper}
\end{document}
\endinput
